---
title: "intRo: Data visualisation with R"
subtitle: "02 — R Basics"
author: "Stefano Coretta"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css:
      - default
      - ../ipa-fonts.css
      - ../custom.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      beforeInit: "../macros.js"
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r xaringanExtra, echo=FALSE}
xaringanExtra::use_xaringan_extra(c("panelset", "tachyons"))
```

# R as a calculator

Write the following in the `Console`, then press `ENTER`.

```{r}
1 + 2
```

--

You did it! You run your first line of code!

---

# R as a calculator

Try some more operations...

```{r}
67 - 13

2 * 4

268 / 43

```

---

# R as a calculator

You can chain operations...

```{r}
6 + 4 - 1 + 2

4 * 2 + 3 * 2
```

---

# R as a calculator

But watch out for your brackets!

```{r}
4 * 2 + 3 * 2

(4 * 2) + (3 * 2)

4 * (2 + 3) * 2
```

---

# Forget-me-not: variables

Store a variable in the computer memory for later use.

```{r}
my_num <- 156
```

--

You can just call it back when you need it!

```{r}
my_num
```

---

# Oh my vars!

```{r}
income <- 1200

expenses <- 500

income - expenses
```

--

You can go all the way with variables!

```{r}
savings <- income - expenses
```

And check the value...

```{r}
savings
```

---

# Variables are... variable

```{r}
yo <- 5

yo + 3

yo <- 3

yo + 3
```

---

# Variables can hold more than one item

```{r}
one_i <- 6

two_i <- c(6, 8)

five_i <- c(6, 8, 42)
```

--

Note the following are the same:

```{r}
one_i <- 6

one_ii <- c(6)
```


---

# R can't function without... functions

**Functions** are operations performed on one or more *arguments*.

```{r}
sum(3, 5)
```

--

And arguments can be vectors!

```{r}
my_nums <- c(3, 5, 7)

sum(my_nums)
mean(my_nums)
```

---

# Not just numbers

Variables can store **strings**.

```{r}
name <- "Stefano"
surname <- "Coretta"

name
```

---

# Not just numbers

And strings can be used as arguments in functions.

```{r}
cat("My name is", name, surname)
```

--

And you can of course reuse the same variable name to override the variable value.

```{r}
name <- "Raj"

cat("My name is", name)
```

---

# The null and the infinite

```{r}
forgot <- NA

nothing <- NULL

everything <- Inf
```


---

# Summary


- R as a calculator

  - Arithmetic operations.

- Variable:
  
  - Numeric.
  - String.

- Function:
  
  - Function name (`sum()`, `mean()`, `cat()`).
  - Arguments.

---

class: middle center inverse

.f1[TUTORIAL]
