---
title: "intRo: Data visualisation with R"
subtitle: "03 — Plot with ggplot2"
author: "Stefano Coretta"
# institute: "RStudio, PBC"
# date: "2016/12/12 (updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

background-image: url(https://upload.wikimedia.org/wikipedia/commons/b/be/Sharingan_triple.svg)

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

???

Image credit: [Wikimedia Commons](https://commons.wikimedia.org/wiki/File:Sharingan_triple.svg)

---

# Visualise data


---

# Graphic systems

- Base R.

- lattice.

- ggplot2 (the Grammar of Graphics).

---

# Base R plotting function

```{r}
x <- 1:10
y <- x^3

plot(x, y)
```

---

# Base R plotting function

```{r}
x <- seq(1, 10, by = 0.01)
y <- x^3

plot(x, y, type = "l", col = "purple", lwd = 3, lty = "dashed")
```

---

# Packages to the rescue

- The library.

- Packages contained in library.

- Base R (set of packages).

- Install extra packages with `install.packages("tidyverse")`.

---

# Packages to the rescue

To use a package installed in the library, you `attach` the package with `library()`.

---

# Packages to the rescue

```{r}
library(tidyverse)
```

---

# Do you sleep like a giraffe or a brown bat?

```{r}
ggplot(
  data = msleep,
  mapping = aes(x = sleep_total, y = sleep_rem)
) +
  geom_point()
```


---

# Do you sleep like a giraffe or a brown bat?


```{r}
ggplot(
  data = msleep,
  mapping = aes(x = sleep_total, y = sleep_rem, label = name)
) +
  geom_point() +
  geom_label()
```

